Racial and Ethnic Differences in ETI: Outcomes

### 2025-08-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
air formatting

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(broom)
library(emmeans)
knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
```

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
# Change age group reference level
encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
```

@@

```{r}
t1_participant %>%
  select(-eDWID) %>%
  tbl_summary(
    by = Race,
    type = list(`Years With Data` ~ "continuous"),
    statistic = list(c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"),
    missing_text = "Missing"
  ) %>%
  add_overall()
```

@@

```{r}
encounter %>%
  select(Race, encounterlocation, bacterialculturedone:othermicroorganisms1) %>%
  tbl_summary(by = Race, missing_text = "Missing")
```

@@

```{r}
# Create change in slope variable
encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_change_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_change_years) <- "Change in Slope"
# Prior modulator use
prior_mods <- encounter %>%
  select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
  filter(Days < 0) %>%
  group_by(eDWID) %>%
  summarise(prior_modulator = any(vx661comb == 1) | any(vx445comb == 1) |
    any(Vx770 == 1) | any(Vx809comb))
prior_mods$prior_modulator <-
  fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
prior_mods$prior_modulator <- factor(prior_mods$prior_modulator,
  levels = c(T, "No"),
  labels = c("Yes", "No")
)
encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
# On enzymes for pancreatic insufficiency
encounter$isOnEnzymes <-
  factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
# Pseudomonas
encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
encounter$pseudomonasaeruginosa <- factor(encounter$pseudomonasaeruginosa,
  levels = 0:1, labels = c("No", "Yes")
)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
  Insurance = "Private", prior_modulator = "Yes",
  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
))
pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(intercept = T) %>%
  as_gt() %>%
  fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt(caption = "Predicted Values at Various Timepoints") %>%
  fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_change,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_change,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_change,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_change
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
    age_group + Insurance + prior_modulator + isOnEnzymes +
    pseudomonasaeruginosa + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(intercept = T) %>%
  as_gt() %>%
  fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_change,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_change,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_change,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_change
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>%
  gtsummary::tbl_regression(intercept = T) %>%
  as_gt() %>%
  fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_change,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_change,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_change,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_change
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_change <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
  "Black/AA Pre-ETI" = aa_pre,
  "Black/AA Post-ETI" = aa_pre + aa_change,
  "H/L Pre-ETI" = hl_pre,
  "H/L Post-ETI" = hl_pre + hl_change,
  "NHW Pre-ETI" = nhw_pre,
  "NHW Post-ETI" = nhw_pre + nhw_change,
  "Other Pre-ETI" = other_pre,
  "Other Post-ETI" = other_pre + other_change
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
  as.data.frame() %>%
  gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

>>>>

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(broom)
library(emmeans)
home_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
    "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Windows" = "C:/Users/Tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load(
    "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData"
)
# Change age group reference level
encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
```

@@

```{r}
t1_participant %>%
    select(-eDWID) %>%
    tbl_summary(
        by = Race,
        type = list(`Years With Data` ~ "continuous"),
        statistic = list(
            c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"
        ),
        missing_text = "Missing"
    ) %>%
    add_overall()
```

@@

```{r}
encounter %>%
    select(
        Race,
        encounterlocation,
        bacterialculturedone:othermicroorganisms1
    ) %>%
    tbl_summary(by = Race, missing_text = "Missing")
```

@@

```{r}
# Create change in slope variable
encounter$slope_c_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_c_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_c_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_c_years) <- "Change in Slope"
# Prior modulator use
prior_mods <- encounter %>%
    select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
    filter(Days < 0) %>%
    group_by(eDWID) %>%
    summarise(
        prior_modulator = any(vx661comb == 1) |
            any(vx445comb == 1) |
            any(Vx770 == 1) |
            any(Vx809comb)
    )
prior_mods$prior_modulator <-
    fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
prior_mods$prior_modulator <- factor(
    prior_mods$prior_modulator,
    levels = c(T, "No"),
    labels = c("Yes", "No")
)
encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
# On enzymes for pancreatic insufficiency
encounter$isOnEnzymes <-
    factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
# Pseudomonas
encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
encounter$pseudomonasaeruginosa <- factor(
    encounter$pseudomonasaeruginosa,
    levels = 0:1,
    labels = c("No", "Yes")
)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
    Years = -3:2,
    Race = unique(encounter$Race),
    age_group = "[18,30)",
    Insurance = "Private",
    prior_modulator = "Yes",
    isOnEnzymes = "Yes",
    pseudomonasaeruginosa = "No"
))
pop_dat$slope_c_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("FEV1 % Predicted") +
    theme_classic(base_size = 15)
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt(caption = "Predicted Values at Various Timepoints") %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```
